---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
con_name <- as.numeric(Sys.time() + 1)
system(glue::glue("docker run --publish=7474:7474 --publish=7687:7687 --env NEO4J_AUTH=neo4j/pouetpouet --name {con_name} neo4j4test"), wait = FALSE)
con <- neo4r::neo4j_api$new(url = "http://localhost:7474",
                            user = "neo4j", password = "pouetpouet")
while(con$ping() != 200) {
  Sys.sleep(1)
}
library(neo4r)
con <- neo4j_api$new(url = "http://localhost:7474", user = "neo4j", password = "pouetpouet")
send_cypher("data-raw/constraints.cypher", con, include_stats = TRUE)


'USING PERIODIC COMMIT 500
LOAD CSV WITH HEADERS FROM "file:///pkg.csv" AS csv 
MERGE (p:Package {name: csv.Package, version: csv.Version, license: csv.License, date: csv.Date, title: csv.Title})
MERGE (m:Maintainer {name: csv.Maintainer})
CREATE (m) -[:MAINTAINS]-> (p);' %>%
  call_api(con, include_stats = TRUE)

'USING PERIODIC COMMIT 500
LOAD CSV WITH HEADERS FROM "file:///depends.csv" AS csv 
MATCH (p:Package {name : csv.Package})
MATCH (l:Package {name: csv.Depends})
CREATE (p) -[:DEPENDS_ON]-> (l);' %>%
  call_api(con, include_stats = TRUE)

'USING PERIODIC COMMIT 500
LOAD CSV WITH HEADERS FROM "file:///imports.csv" AS csv 
MATCH (p:Package {name : csv.Package, version: csv.Version})
MATCH (l:Package {name: csv.Depends})
CREATE (p) -[:IMPORTS]-> (l);' %>%
  call_api(con, include_stats = TRUE)

```

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

> Disclaimer: this package is still at an experimental level and under active development. You should only use it for testing, reporting bugs (which are to be expected), proposing changes to the code, requesting features, sending ideas... As long as this package is in "Experimental" mode, there might be bugs, and changes to the API are to be expected. Read the [NEWS.md](NEWS.md) to be informed of the last changes.


# neo4r

The goal of {neo4r} is to provide a modern and flexible Neo4J driver for R.

It's modern in the sense that you results are returned as tibbles whenever possible, relies on modern tools, and is designed to work with the pipe. Our goal with this package is to provide a driver that can be easily integrated in a data analysis workflow, especially by providing an API that can work smoothly with other data analysis (dplyr or purrr) and graph packages (igraph, ggraph, visNetwork...).

It's flexible in the sense that it is rather unopinionated regarding the way it returns the results, by trying to stay as close as possible to the way Neo4J returns data. That way, you have the control over the way you will compute the results. At the same time, the result is not too complex, so that the "heavy lifting" of data wrangling is not left to the user. 

The connexion object is also an easy to control R6 method, allowing you to update and query information from the API. 

## Installation

You can install {neo4r} from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("neo4j-rstats/neo4r")
```

## Create a connexion object 

Start by creating a new connexion object with `neo4j_api$new`

```{r}
library(neo4r)
con <- neo4j_api$new(url = "http://localhost:7474", 
                     user = "plop", password = "pouetpouet")
```

This connexion object is designed to interact with you Neo4J API. 

It comes with some methods to retrieve information from it : 

```{r}
# Test the endpoint, that will not work :
con$ping()
```

Being an R6 object, `con` is flexible in the sense that you can change `url`, `user` and `password` at any time:

```{r}
con$reset_user("neo4j")
con$ping()
# Or with 
con$password <- "pouetpouet"
```

That means you can at any time connect to another url without having to generate a new connexion object. (`con$reset_url()`).

```{r}
# Get Neo4J Version
con$get_version()
# List constaints (if any)
con$get_constraints()
# Get a vector of labels (if any)
con$get_labels()
# Get a vector of relationships (if any)
con$get_relationships()
# Get schema 
con$get_schema()
```

## Call the API

You can either create a separate query or insert it inside the `call_api` function. 

The `call_api` function takes several arguments : 

+ query : the cypher query
+ con : the connexion object 
+ type : "rows" or "graph": wether to return the results as a list of results in tibble, or as a graph object (with `$nodes` and `$relationships`)
+ output : the output format (r or json) 
+ include_stats : whether or not to include the stats about the call 
+ meta : wether or not to include the meta arguments of the nodes when calling with "rows"

> At the end of the developping process of all the packages planned, you will be able to write your queries and pipe them with {cyphersugar}, which offers a syntactic sugar on top of cypher.

### "rows" format

When you're calling the API, you can choose to returns a list of tibbles. You get as many objects as you specified in the RETURN cypher statement. 

```{r}
library(magrittr)

'MATCH (p:Package) RETURN p.name AS nom LIMIT 5' %>%
  call_api(con)
```

By default, results are returned as an R list of tibbles. We choose to implement it this way, as we think this is the more "truthful" regarding the way you call Neo4J. 

For example, when you want to return two nodes, you'll get two results, in the form of two tibbles (p.name and dep.name here): 

```{r}
'MATCH (p:Package) <-[:DEPENDS_ON]-(dep:Package) RETURN p.name AS nom, dep.name AS maintainer LIMIT 5' %>%
  call_api(con)
```

The result is a two elements list with each element being labelled as what you specified in the cypher query. 

Results can also be returned in JSON:

```{r}
'MATCH (p:Package) <-[:DEPENDS_ON]-(dep:Package) RETURN p.name AS nom, dep.name AS maintainer LIMIT 1' %>%
  call_api(con, output = "json")
```

If you turn the `type` argument to `graph`, you'll get a graph result: 

```{r}
'MATCH p=()-[r:MAINTAINS]->() RETURN p LIMIT 5' %>%
  call_api(con, type = "graph")
```

The result is returned as one node or relationship by row. 

Due to the specific data format of Neo4J, there can be more than one label and propertie by node and relationship. That's why the results is returned, by design, as a list-data.frame. 

We have designed several functions to unnest this : 

+ #### unnest_nodes, that can unnest a node dataframe : 

```{r}
res <- 'MATCH p=()-[r:MAINTAINS]->() RETURN p LIMIT 5' %>%
  call_api(con, type = "graph")
unnest_nodes(res$nodes)
```

Note that this will return NA for the properties not in a node. For example here, we have no 'licence' information for the Maintainer node (that makes sense). 

On the long run, and this is not {neo4r} specific by Neo4J related, a good practice is to have a "name" propertie on each node, so this column will be full here.

You can also either unnest only the properties or the labels : 

```{r}
res$nodes %>%
  unnest_nodes(what = "properties")
```

```{r}
res$nodes %>%
  unnest_nodes(what = "label")
```

+ `unnest_relationships`

There is only one nested column in the relationship table, so the function is quite straightforward : 

```{r}
unnest_relationships(res$relationships)
```

+ `unnest_graph`

This function takes a graph results, and does `unnest_nodes` and `unnest_relationships`. 

```{r}
unnest_graph(res)
```

## Convert for common graph packages

Unless otherwise specified, the functions do an `unnest_graph` before being transformed to a graph object.

### {igraph}

To be converted to a graph object, 

+ the nodes need an id, and a name. By defaut, the node name is assumed to be found in the "name" property return by the graph, but you can specify any other column. The "label" column from Neo4J is renamed "group".

+ relationships needs a start and end, which are startNode and endNode in the Neo4J results. 

```{r}
res %>%
  convert_to("igraph")
```

Which means that you can : 

```{r}
'MATCH p=()-[r:MAINTAINS]->() RETURN p LIMIT 5' %>%
  call_api(con, type = "graph") %>% 
  convert_to("igraph") %>%
  plot()
```

This can also be used with ggraph :

```{r}
library(ggraph)
'MATCH p=()-[r:MAINTAINS]->() RETURN p LIMIT 10' %>%
  call_api(con, type = "graph") %>% 
  convert_to("igraph") %>%
  ggraph() + 
  geom_node_label(aes(label = name, color = group)) +
  geom_edge_link() + 
  theme_graph()
```

### {visNetwork}

```{r eval = FALSE}
network <- 'MATCH p=()-[r:MAINTAINS]->() RETURN p LIMIT 10' %>%
  call_api(con, type = "graph") %>% 
  convert_to("visNetwork")
visNetwork::visNetwork(network$nodes, network$relationships)
```


## Sending data to the api 

You can simply send queries has we have just seen, by writing the cypher query and call the api.  

### Sending an R data.frame

+ `as_nodes` turns a dataframe into a series of nodes : 

```{r message = FALSE}
call_api("CREATE CONSTRAINT ON (a:setosa) ASSERT a.Species IS UNIQUE;", con)
library(dplyr)
iris %>%
  filter(Species == "setosa") %>%
   as_nodes(label = Species) %>%
   call_api(con, include_stats = TRUE)
```

```{r}
'MATCH (n:setosa) RETURN COUNT(*) AS setosa' %>%
   call_api(con)
```

For now, big dataframes can result in API crash if you try to send too much data. We are working on it, so for now you should split you data.frame into pieces (imagine it as doing a `USING PERIODIC COMMIT`).

+ `as_relationships`

// Coming soon

### Reading and sending a cypher file :

+ `read_cypher` reads a cypher file and returns a tibble of all the calls

```{r}
read_cypher("data-raw/create.cypher")
```

+ `send_cypher` reads a cypher file, and send it the the API. By default, the stats are returned. 

```{r}
send_cypher("data-raw/constraints.cypher", con)
```


### Sending csv dataframe to Neo4J

The `load_csv_with_headers` sends an csv from an url to the Neo4J browser. 

The args are : 

+ on_load : the code to execute on load 
+ con : the connexion object 
+ url : the url of the csv to send
+ header : wether or not the csv has a header
+ periodic_commit : the volume for PERIODIC COMMIT
+ as : the AS argument for LOAD CSV
+ format : the format of the result 
+ include_stats : whether or not to include the stats 
+ meta : whether or not to return the meta information

```{r}
# Create the constraints
call_api("CREATE CONSTRAINT ON (a:artist) ASSERT a.name IS UNIQUE;", con)
call_api("CREATE CONSTRAINT ON (al:album) ASSERT al.name IS UNIQUE;", con)
# List constaints (if any)
con$get_constraints()
# Create the query that will create the nodes and relationships
on_load_query <- 'MERGE (a:artist { name: csvLine.artist})
MERGE (al:album {name: csvLine.album_name})
MERGE (a) -[:has_recorded] -> (al)  
RETURN a AS artists, al AS albums;'
# Send the csv 
load_csv(url = "https://raw.githubusercontent.com/ThinkR-open/datasets/master/tracks.csv", 
         con = con, header = TRUE, periodic_commit = 50, 
         as = "csvLine", on_load = on_load_query)
```


Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.

```{r include = FALSE}
system(glue::glue("docker stop {con_name}"))
```

